---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

`rbcpingarch` is an R package providing code to simulate from and fit the Bivariate Conditional Poisson INGARCH(1,1) process proposed in [Piancastelli, Barreto-Souza and Ombao](https://arxiv.org/abs/2011.08799). Our package links to `rstan`, so proper setup of C++ compiler is in order for usage. We recommend testing this via installation of `Rcpp` and `RcppArmadillo`. Once this is done, the following command line installs `rbcpingarch` from its GitHub repository.

``` r
# install.packages("devtools")
devtools::install_github("luizapiancastelli/rbcpingarch")
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rbcpingarch)
library(ggplot2)
library(tidyr)
```

### BCP-INGARCH simulation


The next chunk of code exemplifies using the `rBCPINGARCH` for simulating a realisation of the BCP-INGARCH(1,1) process. This function takes as input the model parameters with `A` being a (2x2) matrix of parameters related to the previous conditional means, `B` a (2x2) matrix of INGARCH parameters on the previous counts, `omega` the intercept vector, `phi` the correlation parameter and `n`, and the trajectory lenght. A brief description of these arguments is included in a call to `help(rBCPINGARCH)`, and a full model specification is in Definition 2.1 of the paper's preprint linked above. 

The output of `rBCPINGARCH` is a (`n`x 2) integer matrix, with one realisation illustrated in the code below. 

```{r, fig.width=6, fig.height=4}
A = diag(c(0.3, 0.1)) #Diagonal (2x2) matrix
B = diag(c(0.2, 0.3)) #Diagonal (2x2) matrix, can be non diagonal as long as stationarity conditions satisfied
omega = c(2,1) 
phi = 0.3 
n = 500  #Time series length

Y = rBCPINGARCH(A, B, omega, phi, n)
head(Y)
cor(Y)

Y_df= data.frame('y1' =Y[,1], 'y2' = Y[,2], 't' = 1:nrow(Y))
Y_df = pivot_longer(Y_df, cols = starts_with('y'), names_to = 'serie', values_to = 'count')

ggplot(Y_df, aes(x = t, y = count, color = serie))+
  geom_line()+theme_bw()+
  labs(y = 'Count', x = 'Time')
```


### Model fit

The bivariate time series of confirmed viral hepatites counts in the nearby Brazilian cities of Goiania and Brasilia  explored in [Piancastelli, Barreto-Souza and Ombao](https://arxiv.org/abs/2011.08799) is originally collected from the open [DATASUS platform](https://datasus.saude.gov.br), included in `rBCPINGARCH` under `data("Y_hep")`. As described in Section 6 of the manuscript, the study period of this version of the data is from 2001 to 2018.

This data is used in the code chunk below to provide guidelines on using the `fit_BCP_INGARCH` to fit a BCP-INGARCH(1,1) process via conditional maximum likelihood estimation. Its arguments are `Y` the observed bivariate count data, and boolean values `A.diag`, `B.diag` specifying the configuration of the `A` and `B` matrices as diagonal (`TRUE`) or non-diagonal (`FALSE`). The diagonal and non-diagonal model fits analysed in the original manuscript are reproduced in the code below.

```{r}
data("Y_hep") #Loads the Brazilian hepatites count data set

fit_diag = fit_BCP_INGARCH(Y_hep, A.diag = TRUE, B.diag = TRUE) #Diagonal model
fit_diag$par #maximised parameter values
fit_diag$se  #asymptotic standard errors
fit_diag$loglik 

fit_nd = fit_BCP_INGARCH(Y_hep, A.diag = TRUE, B.diag = FALSE) #Non-diagonal model
fit_nd$par
fit_nd$se
fit_nd$loglik
```

The output is a named list containing the estimated model parameters `fit$par`, their asymptotic standard  errors `fit$se`, maximed log-likelihood value `fit$loglik`, the fitted process `fit$lambda`, and the initial values used by the algorithm `fit$initial_values`, among others.

Selecting between the nested diagonal and non-diagonal model fits can be done via information criterian with the `model_information` function. This returns the AIC and BIC of the alternative models, both indicating the diagonal case for the hepatites count data.

```{r}
model_information(fit_diag)
model_information(fit_nd)
```

